\documentclass{ximeraXloud}

\title{Terminology To Know}
\begin{document}
\begin{abstract}
    These are important terms and notations for this section.
\end{abstract}
\maketitle

\begin{definition}[Mathematical Reasoning]
    A technique of problem solving that relies on deductive reasoning and symbolic manipulation.
\end{definition}

\begin{definition}[Deductive Reasoning]
    A form of inference that uses a set of given information, and extracts new information that \emph{must} be true as a result. \\For example (eg): Given the information:
    \begin{itemize}
        \item It has rained 
        \item There is no cover for the ground 
    \end{itemize}
    Then we can deduce that the ground is wet. This would be a logical deduction, or ``deductive reasoning".
\end{definition}

\begin{definition}[Information]
    A collection of knowledge or observations used to problem solve.
\end{definition}

\begin{definition}[Data]
    Quantified versions of information.
\end{definition}

\begin{definition}[Model]
    A (collection of) mathematical expressions used to represent a specific situation.
\end{definition}

\begin{definition}[Generalize]
    A process of going from a (somewhat) specific problem (eg a numeric model) to a methodology or (Generalized) Model to attain answers to (less specific) variations of the original problem.
\end{definition}

\begin{definition}[(Mathematical) Symbol]
    A mathematical symbol (usually referred to simply as `symbol' when context is clear) is a non-numeric piece of writing (typically a letter) that holds the place of some (typically numeric) value, to be decided (or deduced) later.
    
        \emph{The conceptual definition (Not to be used as the actual definition)}: A symbol is a letter, in some alphabet, that is a temporary stand-in for a piece of information. It is typically either a variable or replaced by a concrete value when one forms a model.
        
        \textbf{Author's Note:} This is included as we use the term 'symbol' many times throughout this section and thus it should have a formal definition. \textbf{However this term will not be tested on in discussion quizzes or on exams.} It is necessarily abstract given it's usage, but the abstraction is so deep as to be beyond the scope of this course.
\end{definition}

\begin{definition}[Variable]
    A symbol used to represent an unknown piece of data.
\end{definition}

\begin{definition}[Independant Variable]
    A variable whose value is unspecified, but will be provided by the statement of the problem.
\end{definition}

\begin{definition}[Dependent Variable]
    A variable whose value is unknown and is generated by the Model. In most cases dependent variables are determined (solved for) by the Model (See Arbitrary Constant Below).
\end{definition}

\begin{definition}[Arbitrary Constant]
    A symbol that is unaffected by independent variables, and whose value is determined by the model and initial conditions of the system.
\end{definition}

\end{document}