\documentclass{ximeraXloud}

\title{Quadratic Formula}
\begin{document}
\begin{abstract}
    Exploring the usefulness and (mostly) non-usefulness of the quadratic formula
\end{abstract}
\maketitle

\subsection{Quadratic Formula; the least useful tool of the topic!}

    Generally, most students try to memorize the quadratic formula and ignore most of the other content in this section on quadratics for the simple fact that the quadratic formula \textit{always} gives the zeros of a quadratic whereas the other methods often are considerably more work (at least initially when learning them) and don't always yield a useful answer. Nonetheless the quadratic formula is, by far, the least useful thing in this section for the simple fact that it \textit{only works on quadratic forms where you need the zeros}. 
    
    Nonetheless, there are times when you need exactly this information, and when factoring is not a viable option. In these cases, the quadratic formula is quite helpful, and thus you should think of this as a niche tool; a formula that has a very specific circumstance in which it can and should be used, but one which is has very limited scope of usefulness... not the ubiquitous hammer which always yields the correct answer if you just smack the problem enough with it.

    \subsection*{Ok, very specific tool, only use when necessary. Got it... but what is it again? $-b$ something or other?}
    
        The quadratic formula is to find \textit{zeros} of a quadratic, and thus should be written as the full equation (in order to remember what you are solving for, which seems often forgotten with this tool). Thus the quadratic formula statement is as follows: 
    
        \begin{theorem}
            For a quadratic of the form $p(x) = ax^2 + bx + c$, (and $a \neq 0$) the zeros ($x$-values) of the quadratic can be computed by the equation
            \[
            x = \dfrac{-b \pm \sqrt{b^2 - 4ac}}{2a}
            \]
        \end{theorem}
        A fair question may be `how do we know that this works?' Which, as you may remember from the endless harping on it that I do, is where the \textit{precision} of mathematics comes in. We shouldn't take on faith that this equation works, instead we should show how we know for certain that this equation works, ie we should \textit{prove} it.
        
        To do this, notice that what we want are the zeros if $p(x)$, which is to say we want the $x$ values so that $p(x) = 0$, ie we want to solve the equation;%
        \footnote{Yes, prepare thyself for the alphabet soup of doom.}
        \[
            0 = ax^2 + bx + c
        \]
        
        As it happens, we can actually solve this, despite the remarkable level of generality, by... you guessed it... completing the square! Thus we can compute the following:
        
        \begin{example}%
            Deriving the Quadratic Formula using Completing the Square. We will go step by step and explain each step as clearly as possible to show how one uses completing the square to go from the general quadratic form $ax^2 + bx + c$ to the quadratic formula that states the zeros of that form are $x = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
            \begin{align*}
                ax^2 + bx + c & = 0  \hspace{2cm} \text{ Starting point}    \\
                x^2 + \frac{b}{a}x + \frac{c}{a} & = 0    \hspace{2cm} \text{Divide out }a \neq 0  \\
                x^2 + \frac{b}{a}x + \left(\left(\frac{b}{2a}\right)^2 - \left(\frac{b}{2a}\right)^2\right) + \frac{c}{a} & = 0  \hspace{2cm} \text{Add and subtract the $h = \left(\frac{\left(\frac{b}{a}\right)}{2}\right)^2 = \left(\frac{b}{2a}\right)^2$ }    \\
                \left(x + \frac{b}{2a}\right)^2 - \left(\frac{b}{2a}\right)^2 + \frac{c}{a} & = 0    \hspace{2cm} \text{Complete the Square} \\
                \left( x + \frac{b}{2a}\right)^2 - \frac{b^2}{4a^2} + \frac{4ac}{4a^2} & = 0  \hspace{2cm} \text{Find a common denominator}\\
                \left(x + \frac{b}{2a}\right)^2 & = \frac{b^2-4ac}{4a^2} \hspace{2cm} \text{ Move constants to the right} \\
                x + \frac{b}{2a} & = \pm \sqrt{\frac{b^2-4ac}{4a^2}} \hspace{2cm} \text{Square root both sides} \\
                x + \frac{b}{2a} & = \pm \frac{\sqrt{b^2-4ac}}{\sqrt{4a^2}} \hspace{2cm} \text{Split the square root} \\
                x + \frac{b}{2a} & = \pm \frac{\sqrt{b^2-4ac}}{2a} \hspace{2cm} \text{Simplify perfect square denominator} \\
                x &= \pm \frac{\sqrt{b^2-4ac}}{2a} -\frac{b}{2a} \hspace{2cm} \text{Move $\frac{b}{2a}$ from left to right side} \\
                x &= \frac{-b \pm \sqrt{b^2-4ac}}{2a} \hspace{2cm} \text{Combine terms} \\
            \end{align*}
        \end{example}%
        
        Notice that the equation makes no assumptions to suggest that the expression $b^2 - 4ac$ is non-negative. In fact, we can learn a lot thanks to the value of this expression, which also gets a special name; the \textit{discriminate}. In general, the discriminate is some sort of `distance' between the vertex and the zeros of the function. Thus if the discriminate is positive (noting as well that a quadratic is symmetric across the vertex since it's a `U' shape), then it will have 2 (real-valued) zeros, one on each side of the vertex. If it is zero, then it will have the zeros \textit{at} the vertex, meaning the two roots are duplicated, so there is only one unique (real-valued) root with multiplicity 2. Finally, if the discriminate is negative, this means that the `distance' to the zeros is imaginary, which can only happen if there \textit{are no} real-valued zeros. Since we know the roots exist by the fundamental theorem of algebra, they must be non-real valued (ie complex roots). To put it succinctly:
        
        \begin{itemize}
            \item If $b^2 - 4ac > 0$ there are 2 real-valued zeros and thus 2 real-valued roots.
            \item If $b^2 - 4ac = 0$ there is 1 real-valued zero and thus 1 real-valued root with multiplicity 2 or a `double root'.
            \item If $b^2 - 4ac < 0$ there are 0 real-valued zeros and thus 2 complex-valued (ie non-real) roots.
        \end{itemize}


\end{document}